---
title: "BIO593 Assignment 3"
author: "Kevin Wong, Danielle Jordan, Roxy Roja"
date: '2018-03-26'
output: rmarkdown::github_document
---

```{r Intro, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
chooseCRANmirror(graphics=FALSE, ind=1)
```

## Introduction of the data

This document uses data with life-history, population, and ecological examples. This data can be cited here: http://dx.doi.org/10.6084/m9.figshare.1314459. We will provide an indepth analysis containing plots that visually represent trends in the data and statistical analyses to accurately describe the trends. We will outline the major questions with this dataset.

## Introduction of the problem

Our analyses investigate how differential plot types can influences rodent species weight, sex ratios, hindfoot length, and abundances. The following markdown will couple illustrative figures with statistical analyses to determine if plot types affect rodent species. 

```{r installpackages, echo=FALSE, results=FALSE, message=FALSE, warning=FALSE}
#install.packages(c("dbplyr","RSQLite","tidyverse"))
library(dbplyr)
library(RSQLite)
library(tidyverse)
```

```{r loaddata, echo=FALSE, cache=FALSE, results=FALSE, message=FALSE, warning=FALSE}
download.file("https://ndownloader.figshare.com/files/2292169" , "data/portal_data_joined.csv") 
surveys <- read.csv("../data/portal_data_joined.csv", header=TRUE, na.strings = c("","NA")) #uploading csv
```

## Question 1a
Does plot type affect the relationship between hindfoot length and sex of rodent species? 

``` {r Q1a, echo = FALSE}
hyp1 <- hypothesis1 %>% 
  group_by(taxa, sex, plot_type) %>% 
  summarize(avg_weight=mean(weight), sem=sd(weight)/sqrt(length(weight)))

ggplot(data=hyp1, aes(x=plot_type, y=avg_weight)) +
  geom_bar(stat = "identity",aes(fill = sex),
  position = position_dodge(width = 0.9)) +
  ylab("Mean Weight (g)") + xlab("Plot Type")+
  geom_errorbar(aes(ymin=avg_weight+sem, ymax=avg_weight-sem),
                width=0.5,
                size=0.25)
```
## Question 2a
What is the relationship between hindfoot length and species?? 
```{r Q2a, echo = FALSE}
hindspecies <- surveys %>% 
  select(taxa, species_id, hindfoot_length) %>% 
  filter(!is.na(hindfoot_length)) %>% 
  filter(!is.na(species_id)) %>% 
  group_by(taxa,species_id, hindfoot_length)

species_hind <- hindspecies %>% 
  summarize(avg_hind=mean(hindfoot_length), sem=sd(hindfoot_length)/sqrt(length(hindfoot_length))) %>% 
  filter(!is.na(sem))

ggplot(data=species_hind, aes(x=species_id, y=avg_hind))+
  geom_bar(stat="identity",
  position = position_dodge(width = 0.9)) +
  ylab("Average Hindfoot Length (cm)") + xlab("Species ID")+
  geom_errorbar(aes(ymin=avg_hind+sem, ymax=avg_hind-sem),
                width=0.5,
                size=0.25)
```
## Question 2b
What is the relationship between hindfoot length and weight? 
```{r 2b, echo = FALSE}
whind <- weighthind %>% 
  group_by(taxa, species_id, hindfoot_length) %>% 
  summarize(avg_weight=mean(weight), sem=sd(weight)/sqrt(length(weight))) %>% 
  filter(!is.na(sem))

ggplot(data=whind, aes(x=hindfoot_length, y=avg_weight)) +
  geom_bar(stat = "identity",
           position = position_dodge(width = 1.5)) +
  ylab("Mean Weight (g)") + xlab("Hindfoot Length (cm)")+
  geom_errorbar(aes(ymin=avg_weight+sem, ymax=avg_weight-sem),
                width=0.5,
                size=0.25)
```

## Question 2c
Does plot type affect the relationship between weight and sex of rodent species? 
```{r Q2c, echo = FALSE}
hyp1 <- hypothesis1 %>% 
  group_by(taxa, sex, plot_type) %>% 
  summarize(avg_weight=mean(weight), sem=sd(weight)/sqrt(length(weight)))

ggplot(data=hyp1, aes(x=plot_type, y=avg_weight)) +
  geom_bar(stat = "identity",aes(fill = sex),
  position = position_dodge(width = 0.9)) +
  ylab("Mean Weight (g)") + xlab("Plot Type")+
  geom_errorbar(aes(ymin=avg_weight+sem, ymax=avg_weight-sem),
                width=0.5,
                size=0.25)
```

## Question 3a (Plot)
How does rodent species abundance change in each plot type? 
```{r Q3a, echo = FALSE, results = TRUE}
surveys_specplot<- surveys %>% 
  select(taxa, plot_type, species) %>%
  filter(taxa == "Rodent") %>% 
  filter(!is.na(taxa), !is.na(plot_type), !is.na(species)) %>%
  group_by(plot_type, species) %>%
  tally()

ggplot(data = surveys_specplot, aes(x = factor(-n), y = n, fill = species)) +
  geom_bar(stat="identity") +
  ylab("Counts") + xlab("Species") + 
  guides(fill = guide_legend(ncol = 3)) +
  facet_wrap(~ plot_type, scales = "free_x", drop = TRUE) +
  scale_fill_discrete(name="Species") +
  theme_bw(base_size=7) + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  theme(legend.position=c(0.85,0.25)) 

```

## Question 3b (Statistics)
A Pearsons Chi-squared test was used to determine relationship between rodent species abundance and plot type. The residuals are plotted from the Chi-squared analysis to determine postivie and negative associations. Blue means positive association and red means negative association.
```{r Q3b, echo = FALSE, warning=FALSE, message=FALSE, results = TRUE}

tbl = table(surveys$plot_type, surveys$species)

chisqsp <- chisq.test(tbl)
install.packages("devtools")
devtools::install_github("taiyun/corrplot", build_vignettes = TRUE)
install.packages("corrplot")
library(corrplot )

chisqsp

round(chisqsp$residuals, 3)
corrplot(chisqsp$residuals, is.corr=FALSE)
```

## Question 4
How does the total number of rodents change in each plot type over time?
```{r Q4, echo = FALSE, results = TRUE}
surveys_timeplot<- surveys %>% 
  select(taxa, plot_type, species, year) %>%
  filter(taxa == "Rodent") %>% 
  filter(!is.na(taxa), !is.na(plot_type), !is.na(species)) %>%
  group_by(plot_type, year) %>%
  tally()

ggplot(surveys_timeplot,aes(x=year, y=n, color = plot_type))+
  ylab("Total number of rodents")+ xlab("Year") + labs(color ="Plot Type") +
  geom_point()+
  geom_line(aes(group=plot_type))+
  ylim(0,1200) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```