---
title: "BIO593 Assignment 3"
author: "Kevin Wong, Danielle Jordan, Roxy Roja"
date: '2018-03-26'
output: rmarkdown::github_document
---

```{r Intro, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
chooseCRANmirror(graphics=FALSE, ind=1)
```

## Introduction of the data

This document uses data with life-history, population, and ecological examples. This data can be cited here: http://dx.doi.org/10.6084/m9.figshare.1314459. We will provide an indepth analysis containing plots that visually represent trends in the data and statistical analyses to accurately describe the trends. We will outline the major questions with this dataset.

## Introduction of the problem

Our analyses investigate how differential plot types can influences rodent species weight, sex ratios, hindfoot length, and abundances. The following markdown will couple illustrative figures with statistical analyses to determine if plot types affect rodent species. 

```{r installpackages, echo=FALSE, results=FALSE, message=FALSE, warning=FALSE}
#install.packages(c("dbplyr","RSQLite","tidyverse"))
library(dbplyr)
library(RSQLite)
library(tidyverse)
```

```{r loaddata, echo=FALSE, cache=FALSE, results=FALSE, message=FALSE, warning=FALSE}
#download.file("https://ndownloader.figshare.com/files/2292169" , "data/portal_data_joined.csv") 
surveys <- read.csv("../data/portal_data_joined.csv", header=TRUE, na.strings = c("","NA")) #uploading csv
```

## Question 1 (Roxy)
Does plot type affect the relationship between hindfoot length and sex of rodent species? 

``` {r Q1, echo = FALSE}


```
## Question 2a
What is the relationship between hindfoot length and species?? 
```{r Q2a, echo = FALSE}
hindspecies <- surveys %>% 
  select(taxa, species_id, hindfoot_length) %>% 
  filter(!is.na(hindfoot_length)) %>% 
  filter(!is.na(species_id))

species_hind <- hindspecies %>% 
  group_by(species_id) %>% 
  summarize(avg_hind=mean(hindfoot_length), sem=sd(hindfoot_length)/sqrt(length(hindfoot_length))) 

ggplot(data=species_hind, aes(x=species_id, y=avg_hind))+
  geom_bar(stat="identity",
           position = position_dodge(width = 0.9)) +
  ylab("Average Hindfoot Length (cm)") + xlab("Species ID")+
  geom_errorbar(aes(ymin=avg_hind+sem, ymax=avg_hind-sem),position=position_dodge(0.75))
```
## Question 2b
What is the relationship between hindfoot length and weight? 
```{r 2b, echo = FALSE}
weighthind <- surveys %>% 
  select(taxa, species_id, hindfoot_length, weight) %>% 
  filter(!is.na(hindfoot_length)) %>% 
  filter(!is.na(weight))

whind <- weighthind %>% 
  group_by(taxa, species_id, hindfoot_length) %>% 
  summarize(avg_weight=mean(weight), sem=sd(weight)/sqrt(length(weight))) %>% 
  filter(!is.na(sem))

ggplot(data=whind, aes(x=hindfoot_length, y=avg_weight)) +
  geom_point(stat = "identity",
           position = "identity") +
  theme_minimal()+
     ylab("Mean Weight (g)") + xlab("Hindfoot Length (cm)")+
  geom_errorbar(aes(ymin=avg_weight+sem, ymax=avg_weight-sem))
```


## Question 2c 
Does plot type affect the relationship between weight and sex of rodent species? 
```{r Q2c, echo = FALSE}
hypothesis1 <- surveys %>% 
  select(taxa, sex, weight, plot_type) %>%
  filter(!is.na(weight)) %>% 
  filter(!is.na(sex))

hyp1 <- hypothesis1 %>% 
  group_by(taxa, sex, plot_type) %>% 
  summarize(avg_weight=mean(weight), sem=sd(weight)/sqrt(length(weight)))

ggplot(data=hyp1, aes(x=plot_type, y=avg_weight, fill=sex))+
  geom_bar(stat = "identity",aes(fill = sex),
           position = position_dodge(width = 0.9)) +
  ylab("Mean Weight (g)") + xlab("Plot Type")+
  geom_errorbar(aes(ymin=avg_weight+sem, ymax=avg_weight-sem), position=position_dodge(0.9),
                width=0.2,
                size=0.25)
```

## Question 3a (Plot)(Kevin)
How does rodent species abundance change in each plot type? 
```{r Q3a, echo = FALSE, results = TRUE}
surveys_specplot<- surveys %>% 
  select(taxa, plot_type, species) %>% #selecting variables
  filter(taxa == "Rodent") %>% #filtering rodents
  filter(!is.na(taxa), !is.na(plot_type), !is.na(species)) %>% #removing na
  group_by(plot_type, species) %>% #grouping by plot and species
  tally() #getting a count

ggplot(data = surveys_specplot, aes(x = factor(-n), y = n, fill = species)) + #factor allows to re-arrange columns from greatest to smallest
  geom_bar(stat="identity") +
  ylab("Counts") + xlab("Species") + 
  guides(fill = guide_legend(ncol = 3)) + #columns in legend
  facet_wrap(~ plot_type, scales = "free_x", drop = TRUE) + #facet wrap and removing xaxis
  scale_fill_discrete(name="Species") + #labelling legend
  theme_bw(base_size=7) + theme(panel.border = element_blank(), panel.grid.major = element_blank(), #altering legend size and removing grid
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  theme(legend.position=c(0.85,0.25)) #altering legend position

```

## Question 3b (Statistics) (Kevin)
A Pearsons Chi-squared test was used to determine relationship between rodent species abundance and plot type. The residuals are plotted from the Chi-squared analysis to determine postivie and negative associations. Blue means positive association and red means negative association.
```{r Q3b, echo = FALSE, warning=FALSE, message=FALSE, results = TRUE}

tbl = table(surveys$plot_type, surveys$species) #creating a table

chisqsp <- chisq.test(tbl) #Chisqu test
install.packages("devtools")
devtools::install_github("taiyun/corrplot", build_vignettes = TRUE) #downloading corr plot from github
install.packages("corrplot")
library(corrplot )

chisqsp

round(chisqsp$residuals, 3) #getting a table of residulas
corrplot(chisqsp$residuals, is.corr=FALSE) #bubble plot of residulas
```

## Question 4 (Plot)(Kevin)
How does genus abundance differ between plot types?
```{r Q4, echo = FALSE, results=TRUE, warning = FALSE}
surveys_genplot<- surveys %>% 
  select(taxa, plot_type, genus) %>% #selecting variables
  filter(taxa == "Rodent") %>% #Filtering rodent
  filter(!is.na(taxa), !is.na(plot_type), !is.na(genus)) %>% #removing NAs
  group_by(plot_type, genus) %>% #grouping
  tally() #counts

ggplot(data = surveys_genplot, aes(x = factor(-n), y = n, fill = genus)) + #factor allows to re-arrange columsn from greatest to smallest
  geom_bar(stat="identity") +
  ylab("Counts") + xlab("Genus") + 
  guides(fill = guide_legend(ncol = 2)) + #number of columns in legend
  facet_wrap(~ plot_type, scales = "free_x", drop = TRUE) +
  scale_fill_discrete(name="Genus") +
  theme_bw(base_size=8) + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  theme(legend.position=c(0.85,0.25)) 
```

## Question 4b (Statistics)(Kevin)
A Pearsons Chi-squared test was used to determine relationship between rodent genus abundance and plot type. The residuals are plotted from the Chi-squared analysis to determine postivie and negative associations. Blue means positive association and red means negative association.

```{r Q4b, echo = FALSE, results=TRUE, warning=FALSE}
tbl2 = table(surveys$plot_type, surveys$genus)
tbl2

chisqsp2 <- chisq.test(tbl2)

round(chisqsp2$residuals, 3)
corrplot(chisqsp2$residuals, is.corr=FALSE)
```

## Question 5 (Plot)(Kevin)
How does the total number of rodents change in each plot type over time?
```{r Q5, echo = FALSE, results = TRUE}
surveys_timeplot<- surveys %>% 
  select(taxa, plot_type, species, year) %>%
  filter(taxa == "Rodent") %>% 
  filter(!is.na(taxa), !is.na(plot_type), !is.na(species)) %>%
  group_by(plot_type, year) %>%
  tally()

ggplot(surveys_timeplot,aes(x=year, y=n, color = plot_type))+
  ylab("Total number of rodents")+ xlab("Year") + labs(color ="Plot Type") +
  geom_point()+
  geom_line(aes(group=plot_type))+
  ylim(0,1200) + #setting yaxis limits
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Question 5b (Statistics)(Kevin)
Using a Generalized Lease Square model to determine the effects of plot type over year.
```{r Q5b, echo = FALSE, results=TRUE, warning = FALSE}
install.packages("nlme")
library(nlme)

glstime <- gls(n ~ year*plot_type, data = surveys_timeplot)
plot(glstime)
summary(glstime)
```

## Question 6 (Danielle)
How does hindfoot length change with weight?
```{r Q6, echo = FALSE}
weighthind <- surveys %>% 
  select(taxa, species_id, hindfoot_length, weight) %>% 
  filter(!is.na(hindfoot_length)) %>% 
  filter(!is.na(weight))

whind <- weighthind %>% 
  group_by(taxa, species_id, hindfoot_length) %>% 
  summarize(avg_weight=mean(weight), sem=sd(weight)/sqrt(length(weight))) %>% 
  filter(!is.na(sem))

ggplot(data=whind, aes(x=hindfoot_length, y=avg_weight)) +
  geom_bar(stat = "identity",
           position = position_dodge(width = 1.5)) +
  ylab("Mean Weight (g)") + xlab("Hindfoot Length (cm)")+
  geom_errorbar(aes(ymin=avg_weight+sem, ymax=avg_weight-sem),
                width=0.5,
                size=0.25)

```

## Question 7 (Danielle)

How does hindfoot length change with species ID?
```{r Q7, echo = FALSE}
hindspecies <- surveys %>% 
  select(taxa, species_id, hindfoot_length) %>% 
  filter(!is.na(hindfoot_length)) %>% 
  filter(!is.na(species_id)) %>% 
  group_by(taxa,species_id, hindfoot_length)

species_hind <- hindspecies %>% 
  summarize(avg_hind=mean(hindfoot_length), sem=sd(hindfoot_length)/sqrt(length(hindfoot_length))) %>% 
  filter(!is.na(sem))

ggplot(data=species_hind, aes(x=species_id, y=avg_hind))+
  geom_bar(stat="identity",
  position = position_dodge(width = 0.9)) +
  ylab("Average Hindfoot Length (cm)") + xlab("Species ID")+
  geom_errorbar(aes(ymin=avg_hind+sem, ymax=avg_hind-sem),
                width=0.5,
                size=0.25)
```